---
title: "HW2"
output: 
  pdf_document: 
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
1.	Create a new column “Enrollment group” in the table Phonecall
a)	Insert EnrollmentGroup=Clinical Alert :code is 125060000
b)	Insert EnrollmentGroup =Health Coaching :code is 125060001
c)	Insert EnrollmentGroup =Technixal Question: Code is 125060002
d)	Insert EnrollmentGroup =Administrative: Code  is 125060003
e)	Insert EnrollmentGroup =Other: Code  is 125060004
f)	Insert EnrollmentGroup =Lack of engagement : Code  is 125060005

``` {r q1}
library("RODBC")
library(tidyverse)

myconn<-odbcConnect("qbs181","aivanov","aivanov@qbs181")
phonecall<-sqlQuery(myconn,"select * from PhoneCall_Encounter")
library("dplyr")



phonecall <- mutate( phonecall, 'Enrollment group' = ifelse(EncounterCode == '125060000', 'Clinical Alert',
                ifelse(EncounterCode == '125060001', 'Health Coaching',
                 ifelse(EncounterCode == '125060002', 'Technical Question',
                        ifelse(EncounterCode == '125060003', 'Administrative',
                               ifelse(EncounterCode == '125060004', 'Other',
                                      ifelse(EncounterCode == '125060005', 'Lack of engagement',EncounterCode)))))))

sample_n(phonecall, 10)
```

# Question 2 

2.	Obtain the # of records for each enrollment group
              
``` {r q2}
#table(phonecall$`Enrollment group`)

phonecall %>%
  group_by(`Enrollment group`) %>%
  summarize("number of records" = n())

```

# Question 3

3.	Merge the Phone call encounter table with Call duration table.

``` {r q3}
merge_phone <-sqlQuery(myconn,"select A.*, B.* from PhoneCall_Encounter A
inner Join
PhoneCall B
on
A.CustomerId= B.tri_CustomerIDEntityReference")

head(merge_phone)

#JOINING IN R; SLIGHT DIFFERENCE OF 4; I USED the SQL INNER JOIN
#phonecall2 <-sqlQuery(myconn,"select * from PhoneCall")
#p12 <- merge(phonecall, phonecall2, by.x=c('CustomerId'),
#    by.y=c('tri_CustomerIDEntityReference'))

merge_phone <- mutate(merge_phone, 'Enrollment group' = ifelse(EncounterCode == '125060000', 'Clinical Alert',
                ifelse(EncounterCode == '125060001', 'Health Coaching',
                 ifelse(EncounterCode == '125060002', 'Technical Question',
                        ifelse(EncounterCode == '125060003', 'Administrative',
                               ifelse(EncounterCode == '125060004', 'Other',
                                      ifelse(EncounterCode == '125060005', 'Lack of engagement',EncounterCode)))))))

sample_n(merge_phone, 10)

```

# Question 4

4.	Find out the # of records for different call outcomes and call type. Use 1-Inbound and 2-Outbound, for call types; use 1-No response,2-Left voice mail and 3 successful. Please also find the call duration for each of the enrollment groups 


``` {r q4}
#table(merge_phone$CallType)
#table(merge_phone$CallOutcome)


merge_phone %>%
  group_by(`CallType`) %>%
  summarize("number of records" = n())

merge_phone %>%
  group_by(`CallOutcome`) %>%
  summarize("number of records" = n())

merge_phone %>%
  group_by(`Enrollment group`) %>%
  summarize(sum_call = sum(CallDuration), mean_call = mean(CallDuration))
```




# Question 5

5.	Merge the tables Demographics, Conditions and TextMessages. Find the # of texts/per week, by the type of sender. Draw a visual using ggplot to obtain # of texts and color it by the type of sender


``` {r q5}
library(lubridate)

dem_con_text <-sqlQuery(myconn,"Select A.*, B.*, C.* from Demographics A
Inner Join
Conditions B
on
A.contactid = B.tri_patientid
inner join
TextMessages C
on
A.contactid = C.tri_contactid")



head(dem_con_text)

dem_con_text$TextSentDate <- as.Date(dem_con_text$TextSentDate, format="%m/%d/%y")

sample_n(dem_con_text,10)
#finding weeks using lubridate


q5 <- dem_con_text %>% 
  group_by(SenderName, week = floor_date(TextSentDate, unit = "week")) %>% 
  summarize("Number of Texts" = n())




#finding weeks using cut, keeps start of week.
#q5 <- dem_con_text %>% 
#  group_by(SenderName, week = cut(TextSentDate, "week"))%>%
#  summarize("number of texts" = n())
  

ggplot(q5, aes(fill=SenderName, y=`Number of Texts`, x=week)) +
  geom_bar(position="stack", stat="identity") +
  labs(title = "Sender Name Vs. Number of Texts per week")+
   theme(plot.title = element_text(hjust = 0.5))

```


# Question 6

6.	Obtain the count of texts based on the chronic condition over a period of time (say per week). Draw a visual using ggplot to obtain the counts
``` {r q6}

q6 <- dem_con_text %>% 
  group_by("Chronic Condition" = tri_name, week = floor_date(TextSentDate,"week")) %>% 
  summarize("Number of Texts" =n())

ggplot(q6, aes(fill=`Chronic Condition`, y=`Number of Texts`, x=week)) +
  geom_bar(position="stack", stat="identity") +
  labs(title = "Sender Name Vs. Number of Texts per Week") +
  theme(plot.title = element_text(hjust = 0.5))

```